---
title: "SQL"
output: html_document
---

## FILTERING COLUMNS AND ROWS

With SQL, you can filter columns and rows by using SELECT and WHERE. Let's look at an example using the Lahman database.So we should first load that database, along iwth the package sqldf.

```{r message=FALSE, warning=FALSE}
library(Lahman)
library(sqldf)
```

Suppose now I would like to see the homerun totals for the 1927 Yankees. I could write the following query:

```{r}
query<-"select playerID,yearID,teamID,HR from Batting
WHERE teamID='NYA' and yearID=1927"
sqldf(query)
```

# Equal to or more than
Suppose I would like to find an equal to or more than in a list. I could ues this example: All instances where Yankees have hit 40 HR or more.

```{r}
Query<-"Select playerID,yearID,teamID,HR FROM Batting
WHERE teamID='NYA' and HR>=40"
sqldf(Query)
```

# More than and Less than example
If you would like to look at an example with more than's and less than's I could use this example: All instances where a player had more than 40 HR but less than 60 SO.
```{r}
Query<-"Select playerID,yearID,teamID,HR,SO FROM Batting
WHERE HR>40 and SO<60"
sqldf(Query)
```

# More than example
If I want to look at just than more than's, this is an exmaple: All instances of Phillies in the 1970's hitting more than 30 HR

```{r}
Query<-"Select playerID, yearID, teamID, HR FROM Batting 
WHERE teamID='PHI'and yearID>=1970 and yearID<1980 and HR>30"
sqldf(Query)
```


## ORDER BY

With SQL, you can order the rows of your output, For instance, suppose you want to see every instance of a player hitting more than 50 homeruns. But you would like the players with the most homeruns to be at the top. We could do the following:

```{r}
query<-"SELECT playerID, teamID, yearID, HR FROM Batting 
WHERE HR>50 
ORDER BY HR DESC"
sqldf(query)

```


Here is an example of less than, great than or equal to, and order by alll in one. Example: Find all instances oif a player striking out less than 10 times. Making sure each player has had at least 400 AB's and ordering by strikeouts.

```{r}
query<-"Select playerID, yearID, teamID, SO, AB FROM Batting
WHERE SO<10 and AB>=400
ORDER BY SO"
sqldf(query)
```

##AGGREGATION

Aggregating on SQL is helpful to look at sum's or squashing a section together. 
IF I wanted to look at Ruth's HR I could look at this example.
```{r}
query<-"Select playerID, HR FROM Batting
where playerID='ruthba01'"
sqldf(query)
```

However, IF i wanted to aggregate the HR I could use this.
```{r}
query<-"Select playerID,sum(HR) FROM Batting
where playerID='ruthba01'
GROUP BY playerID"
sqldf(query)
```

#SUM example
To find the career HR toals for all players, but limit the output totals to 600 or more and order it by career total, I could use this
```{r}
query<-"SELECT playerID,sum(HR) FROM Batting
GROUP BY playerID
HAVING sum(HR)>=600
ORDER BY sum(HR)"
sqldf(query)
```

When using aggregations you can do more than just the sum. You can search for the average, max, or minimum, basically to squash the numbers down. Here is an example of averaging and also ordering is by the highest to lowest.
```{r}
query<-"SELECT playerID,avg(HR) FROM Batting
GROUP BY playerID
HAVING avg(HR)>30
ORDER BY sum(HR) DESC"
sqldf(query)
```

##INNER JOIN
Suppose I wanted to look at data from the Batting table, but also the Master table. You can use the term "INNER JOIN" to search both tables. Here is an example.(Side note: remember to always avoid ambiguity, with specifics in each table)
```{r}
query<-"SELECT nameFirst, nameLast,teamID,yearID,HR 
FROM Batting INNER JOIN Master
ON Batting.playerID=Master.playerID
WHERE Batting.playerID='ruthba01'"

sqldf(query)
```

Another example of joining two tables would be: "Find all instances of players hitting more than 50 HR including their First Name, Last Name, team, year, and HR"
```{r}
query<-"SELECT nameFirst, nameLast,teamID,yearID,HR 
FROM Batting INNER JOIN Master
ON Batting.playerID=Master.playerID
WHERE HR>50
ORDER BY HR DESC"

sqldf(query)

```


Suppose you want to join two tables and join more than one specific such as teamID and yearID. You could look for B.Ruths playerID, team name, yearID, HR
```{r}
query<-"SELECT playerID, name,Batting.yearID,Batting.HR
FROM Batting INNER JOIN Teams
ON Batting.teamID=Teams.teamID and Batting.yearID=Teams.yearID
WHERE playerID='ruthba01'"

sqldf(query)
```


#More Join's examples
If you were looking for Babe Ruths first and last name playerID, teamname, yearID, HR,  you could use this code
```{r}
query<-"SELECT nameFirst,nameLast,name,Batting.yearID,Batting.HR
FROM (Batting INNER JOIN Master
ON Batting.playerID=Master.playerID) INNER JOIN Teams
ON Batting.teamID=Teams.teamID AND Batting.yearID=Teams.yearID
WHERE Batting.playerID='ruthba01'"

sqldf(query)
```


A "join" example and using sum'syou could look at career homerun totals for players while limiting the display to just those that hit at least 600 HR's; Having HR totals at the top, first and last names, and HR totals.
```{r}
query<-"SELECT nameFirst,nameLast, sum(HR)
FROM Batting INNER JOIN Master
ON Batting.playerID=Master.playerID
GROUP BY Batting.playerID
HAVING sum(HR)>=600
ORDER BY sum(HR) DESC"

sqldf(query)
```

#Salaries joined 

Example where the first name of player is Bob who all averaged making over $1,000,000 in their career, distinguishing them by their last names and in descending order
```{r}
query<-"SELECT nameFirst,nameLast,avg(salary)
FROM Salaries INNER JOIN Master
ON Salaries.playerID=Master.playerID
WHERE nameFirst='Bob'
GROUP by Salaries.playerID
HAVING avg(salary)>1000000
ORDER BY avg(salary) DESC"

sqldf(query)
```



